---
title: "Umap-plots"
author: "Rojashree Jayakumar"
date: "2022-10-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r}
require(Seurat)
require(data.table)
require(tidyverse)
require(ggpubr)
require(cowplot)
```

define specific directories
```{r}
#directory to save the plots
p_dir = file.path("..","Results")

#load some functions - themes

source("Helper_scripts/figure_themes.R")
```


Run in GPU/cluster - do not run this here

There are 5 brain regions and 32 donors (5x32 samples)
This chunck performs r-PCA integration on all the 32 donors for each brain region and outputs the integrated and clustered data
```{r}

#specify paths


data_path <- "/autofs/space/mindds_001/projects/AbbvieSnRNASeq/data/"
scripts_path <- "/autofs/space/mindds_001/projects/AbbvieSnRNASeq/scripts/"
annie_path <- "/autofs/space/mindds_001/projects/AbbvieSnRNASeq/scripts/annie/"
plot_dir <- paste0(annie_path, "results/plots/manuscript/")

# Define paths for data input/output


rdata_path_in <- paste0(data_path, "Rdata/")
rdata_path_out <- paste0(data_path, "Rdata/annie/")
results_path <- paste0(annie_path, "results/")

brain_regions <- c("BA20", "BA46", "EC", "V1", "V2")

# Use 4 cores

ncore <- 4

# Load packages


library(Seurat)
library(data.table)
library(SeuratWrappers)
library(tidyverse)
library(rstatix)
library(ggpubr)
library(cowplot)
library(RColorBrewer)
library(doParallel)
library(foreach)
select <- dplyr::select
theme_set(theme_cowplot())


# function to integrate the data


process_and_integrate_raw_data <- function(seurat.obj, # Input seurat object
                                           brain_region, # Brain region, e.g. BA46
                                           rdata_file_out, # path to output .Rdata file
                                           regress_vars=c("nCount_RNA", "percent.mito"), # variables to regress out
                                           n_HVGs=2000 # Number of highly-variable genes
) {

  # Print the number of genes and cells in the dataset
  cat(sprintf("Original %s data have %s genes.\n", brain_region, nrow(seurat.obj@assays$RNA@data)))
  cat(sprintf("Original %s data have %s cells.\n", brain_region, ncol(seurat.obj@assays$RNA@data)))

  # Split seurat data by donor ID (ADRC #)
  seurat_split <- SplitObject(seurat.obj, split.by="Donor.ID")

  # Free up memory in workspace
  rm(seurat.obj)

  # normalize and identify variable features for each dataset independently
  cat("\nNormalizing data\n")
  seurat_split <- lapply(X = seurat_split, FUN = function(x) {
    x <- NormalizeData(x, verbose=F) # lognormalize
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = n_HVGs,
                              verbose=F) # find top 2,000 variable genes
  })

  # Save point in case r crashes
  save(seurat_split, file=rdata_file_out)

  # select features that are repeatedly variable across datasets for integration run PCA on each
  # dataset using these features
  cat("\nSelecting integration features\n")
  anchor_features <- SelectIntegrationFeatures(object.list = seurat_split, nfeatures = n_HVGs)
  seurat_split <- lapply(X = seurat_split, FUN = function(x) {
    x <- ScaleData(x, features = anchor_features, verbose = F)
    x <- RunPCA(x, features = anchor_features, verbose = F)
  })

  # Save point in case r crashes
  save(seurat_split, anchor_features, file=rdata_file_out)

  # Perform integration using reciprocal PCA (rPCA)
  cat("\nFinding integration anchors\n")
  integration_anchors = FindIntegrationAnchors(object.list = seurat_split,
                                               anchor.features = anchor_features,
                                               reduction = "rpca")

  # Save point in case r crashes
  save(integration_anchors, file=rdata_file_out)

  # this command creates an 'integrated' data assay
  seurat_proc <- IntegrateData(anchorset = integration_anchors)

  # Save point in case r crashes
  save(seurat_proc, file=rdata_file_out)

  # Free up memory in workspace
  rm(seurat_split, integration_anchors)

  # Specify that the next analyses should be on the integrated assay
  DefaultAssay(seurat_proc) <- "integrated"

  # scale integrated data and run regular PCA for dimensionality reduction
  cat("\nScaling integrated data and performing PCA\n")
  seurat_proc <- ScaleData(seurat_proc, verbose = T, vars.to.regress = regress_vars)
  seurat_proc <- RunPCA(seurat_proc, npcs = 50, verbose = F)

  # UMAP nonlinear dimension reduction
  seurat_proc <- RunUMAP(seurat_proc, reduction = "pca", dims = 1:30)

  # Graph-based neighbor detection
  seurat_proc <- FindNeighbors(seurat_proc, reduction = "pca", dims = 1:30)

  # Identify clusters at resolutions 0.1 through 1.0
  seurat_proc <- FindClusters(seurat_proc, resolution = seq(0.1, 1, by=0.1))

  # Assign brain region name to variable and save
  assign(paste0(brain_region, "_rPCA"), seurat_proc)
  save(list=ls(pattern=("_rPCA")), file=rdata_file_out)

  # Clean up workspace
  rm(seurat_proc)

}

# Iterate over our brain regions


for (brain_region in c("BA46", "V1", "V2")) {
  print(sprintf("Now starting %s", brain_region))

  # Define input and output Rdata files
  rdata_in_file <- paste0(rdata_path_in, sprintf("%s_preprocessed.Rdata", brain_region)) # raw data
  rdata_out_file <- paste0(rdata_path_out, sprintf("%s_rPCA_new.Rdata", brain_region)) # integrated data

  if (!file.exists(rdata_out_file)) {
    print(sprintf("Now processing %s", brain_region))

    # Load raw snRNAseq data
    load(rdata_in_file)

    # Run rPCA integration function on data for current brain region
    process_and_integrate_raw_data(seurat.obj=get(brain_region),
                                   brain_region=brain_region,
                                   rdata_file_out=rdata_out_file,
                                   regress_vars=regress_vars,
                                   n_HVGs=2000)
  }
}

```

Next step is to annotate the cell clusters (the scripts here /autofs/space/mindds_001/projects/AbbvieSnRNASeq/scripts/annie/scripts/All_Nuclei_Integration have to be cleaned)

Now plotting the UMAPs and violin plots 
following chuncks of code have custom functions to plot 

a] the umaps of clustering for all the brain regions and all the cells.
b] expression levels of different cell type markers

```{r}
#' custom function for figure 1b
#' @param obj seurat object
#' @param region string; brain region (as figure title)
#' @param show_label T/F; show label or not; default = F
#' @return ggplot object of dim plot

dim_plot = function(obj, region, show_label = T){
  p = DimPlot(obj, size = size, raster = F, label = show_label, group.by = "celltype") +
      umap_theme() +
      scale_color_manual(
        values = c(
          "astrocytes" = "#F38181", 
          "microglia" = "#A3DE83", 
          "endothelial" = "#FCE38A", 
          "neurons" = "#EA99D5", 
          "oligodendrocytes" = "#8BDBF5"
      )) +
    labs(title = region)
  return(p)
}


#' custom function for figure 1c
#' @param obj seurat object
#' @param genes vector; genes in violin plots
#' @return ggplot object

vlnplot = function(obj, genes){
  DefaultAssay(obj) = "RNA"
  obj@active.ident  = factor(
    obj$celltype,
    levels = c("astrocytes", "microglia", "endothelial","neurons", "oligodendrocytes"))
  p = VlnPlot(
    obj, 
    features = genes, 
    stack = TRUE, sort = FALSE, flip = TRUE
  ) +
    scale_fill_manual(
      values = c(
        # astrocytes
        "ADGRV1" = "#F38181", 
        "ALDH1L1" = "#F38181", 
        "AQP4" = "#F38181",
        "GFAP" = "#F38181",
        # microglia
        "CD74" = "#A3DE83", 
        "P2RY12" = "#A3DE83", 
        # endothelial
        "CLDN5" = "#FCE38A", 
        "VWF" = "#FCE38A", 
        # neurons
        "GRIA1" = "#EA99D5", 
        "GRIK2" = "#EA99D5", 
        # oligodendrocytes
        "MOBP" = "#8BDBF5", 
        "PLP1" = "#8BDBF5"
      )) +
    theme(
      legend.position = 'none',
      axis.title.x=element_blank(),
      plot.background = element_rect(
        fill = "white", 
        color = NA
      ))
  return(p)
}
```

Custom function for the violin plots which show expression levels for markers
```{r}
# marker genes
markers = fread(file.path("..","..","/Data/celltype_markers_ASP.csv"))
vlnplot_genes = c("ADGRV1", "ALDH1L1", "AQP4", "GFAP",
                  "CD74", "P2RY12","VWF", "CLDN5")
vlnplot_genes = markers[gene %in% vlnplot_genes, ]
vlnplot_genes[, celltype := factor(celltype, 
      levels = c("Astrocytes", "Microglia", "Endothelial", 
                  "Neurons", "Oligodendrocytes"))]
vlnplot_genes = vlnplot_genes[order(celltype), ]
vlnplot_genes[, gene := factor(gene, levels = vlnplot_genes$gene)]
```


#EC
```{r}
# load data
load(file.path("..", "Data/Fig1_data/EC_subset.RData"))
EC_subset$celltype = factor(
  EC_subset$Annie_Cell_Type, 
  labels = c("astrocytes", "endothelial", "microglia", 
             "neurons", "oligodendrocytes")
)
#Idents(EC_subset) = "celltype"
#EC_subset = EC_subset[, !Idents(EC_subset) %in% c("unknown")]

# umap
p_umap_EC = dim_plot(EC_subset,"EC")

# violin plot
p_vln_EC = vlnplot(
  obj = EC_subset, 
  genes = vlnplot_genes$gene
)

fig1bc_EC = ggarrange(
  p_umap_EC,
  p_vln_EC,
  ncol = 1, nrow = 2,
  heights = c(1, 1.7)
)

# ggsave(fig1bc_EC, filename = "~/Dropbox (Partners HealthCare)/Abbvie collaboration/Astrocytes scripts/Results/Figure_1/figure1b&1c_EC.png", 
#     height = 10, width = 4)
#rm(EC_subset)
```


#ITG
```{r}
# load data
load(file.path("..", "Data/Fig1_data/ITG_subset.RData"))
#Idents(rPCA) = "celltype"
#rPCA = rPCA[, !Idents(rPCA) %in% c("unknown")]

# umap
p_umap_ITG = dim_plot(
  obj = ITG_subset, 
  region = "ITG"
)
# ggsave(p_umap_ITG, filename = "fig1b_BA20.png", width = 4, height = 4)

# violin plot
p_vln_ITG = vlnplot(
  obj = ITG_subset, 
  genes = vlnplot_genes$gene
)
# ggsave(p, filename = file.path(p_dir, "fig1c_ITG.pdf")), height = 7, width = 4)

fig1bc_ITG = ggarrange(
  p_umap_ITG,
  p_vln_ITG,
  ncol = 1, nrow = 2,
  heights = c(1, 1.7)
)
# ggsave(fig1bc_ITG, filename = file.path(p_dir, "figure1b&1c_ITG.png"), 
#       height = 10, width = 4)
#rm(rPCA)
```

## PFC (BA46)

```{r}
# load data
#Note the data loaded here is the subset of data 

load(file.path("..","Data/Fig1_data/BA46_subset.RData"))
BA46_subset$celltype = factor(
  BA46_subset$Annie_Cell_Type, 
  labels = c("astrocytes", "endothelial", "microglia", 
             "neurons", "oligodendrocytes")
)
#Idents(BA46_rPCA) = "celltype"
#BA46_rPCA = BA46_rPCA[, Idents(BA46_rPCA) != "unknown"]

# umap
p_umap_PFC = dim_plot(
  obj = BA46_subset, 
  region = "PFC"
)
# ggsave(p_EC, filename = file.path(p_dir, "fig1b_PFC.png"), width = 4, height = 4)

# violin plot
p_vln_PFC = vlnplot(
  obj = BA46_subset, 
  genes = vlnplot_genes$gene
)
# ggsave(p, filename = file.path(p_dir, "fig_1c_PFC.pdf"), height = 7, width = 4)

fig1bc_PFC = ggarrange(
  p_umap_PFC,
  p_vln_PFC,
  ncol = 1, nrow = 2,
  heights = c(1, 1.7)
)
# ggsave(fig1bc_PFC, filename = "~/Dropbox (Partners HealthCare)/Abbvie collaboration/Astrocytes scripts/Results/Figure_1/figure1b&1c_PFC.png"), 
#        height = 10, width = 4)
#rm(BA46_rPCA)
```

## V2

```{r}
# load data
load(file.path("..","Data/Fig1_data/V2_subset.RData"))
V2_subset$celltype = factor(V2_subset$Annie_Cell_Type, 
       labels = c("astrocytes", "endothelial", "microglia", 
                  "neurons", "oligodendrocytes"))
#Idents(V2_rPCA) = "celltype"
#V2_rPCA = V2_rPCA[, Idents(V2_rPCA) != "unknown"]

# umap
p_umap_V2 = dim_plot(
  obj = V2_subset, 
  region = "V2"
)
# ggsave(p_EC, filename = file.path(p_dir, "fig1b_V2.png"), width = 4, height = 4)

# violin plot
p_vln_V2 = vlnplot(
  obj = V2_subset, 
  genes = vlnplot_genes$gene
)
# ggsave(p, filename = file.path(p_dir, "fig_1c_V2.pdf"), height = 7, width = 4)

fig1bc_V2 = ggarrange(
  p_umap_V2,
  p_vln_V2,
  ncol = 1, nrow = 2,
  heights = c(1, 1.7)
)
# ggsave(fig1bc_V2, filename = "~/Dropbox (Partners HealthCare)/Abbvie collaboration/Astrocytes scripts/Results/Figure_1/figure1b&1c_V2.png"), 
#        height = 10, width = 4)
#rm(V2_rPCA)
```

## V1

```{r}
load(file.path("..","Data/Fig1_data/V1_subset.RData"))
V1_subset$celltype = factor(V1_subset$Annie_Cell_Type, 
       labels = c("astrocytes", "endothelial", "microglia", 
                  "neurons", "oligodendrocytes","unknown"))
Idents(V1_subset) = "celltype"
V1_subset = V1_subset[, Idents(V1_subset) != "unknown"]

# umap
p_umap_V1 = dim_plot(
  obj = V1_subset, 
  region = "V1"
)
# ggsave(p_EC, filename = file.path(p_dir, "fig1b_V1.png"), width = 4, height = 4)

# violin plot
p_vln_V1 = vlnplot(
  obj = V1_subset, 
  genes = vlnplot_genes$gene
)
# ggsave(p, filename = file.path(p_dir, "fig_1c_V1.pdf"), height = 7, width = 4)

fig1bc_V1 = ggarrange(
  p_umap_V1,
  p_vln_V1,
  ncol = 1, nrow = 2,
  heights = c(1, 1.7)
)
#ggsave(fig1bc_V1, filename = "~/Dropbox (Partners HealthCare)/Abbvie collaboration/Astrocytes scripts/Results/Figure_1/figure1b&1c_V1.png"), 
#       height = 10, width = 4)
#rm(V1_rPCA)
```

