---
title: "Correlation analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(
  root.dir = "~/Dropbox (Partners HealthCare)/Abbvie collaboration/Astrocytes scripts"
)
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE, 
  message = FALSE
) 

# load required packages
require(Seurat)
require(data.table)
require(tidyverse)
require(circlize)
require(ComplexHeatmap)
data_dir = file.path("..", "Data")
rdir = file.path("..", "Results")
getwd()

# directories

```

```{r}
p_dir = file.path("..", "Results")
source("Scripts/Helper_scripts/figure_themes.R")
```

Merged data
```{r}
# load astrocytes data
#load(file.path("Data/Fig5_data/cluster_groups_subset.RData"))

#To get the astrocyte data, run the following code on mindds

# directories
data_dir = file.path("..", "..", "data")
results_dir = file.path("..", "..", "results")
p_dir = file.path(results_dir, "plots", "all_brain_regions")

# load astrocyte integrated data
# BA20 ast
load(file.path(data_dir, "Rdata", "ast_removed_logNormal_CCA_R4.0.Rdata"))
BA20_ast = ast_CCA
rm(ast_CCA)

# BA46 ast
load(file.path(data_dir, "BA46_ast_CCA_data1B_0614.Rdata"))
BA46_ast = ast_CCA
rm(ast_CCA)

# V2 ast
load(file.path(data_dir, "V2_ast_CCA_data1B_0614.Rdata"))
V2_ast = ast_CCA
rm(ast_CCA)

# V1 ast
load(file.path(data_dir, "V1_ast_CCA_data1B_0614.Rdata"))
V1_ast = ast_CCA
rm(ast_CCA)

# EC ast
load(file.path(data_dir, "EC_ast_CCA_data1B_0614.Rdata"))
EC_ast = ast_CCA
rm(ast_CCA)

# load cluster annotation
clust_annot2 = fread(file.path(results_dir, "ast_cluster.annotation_HVG.after_Dec08.csv"), 
                     stringsAsFactors = F)
clust_annot2[, cluster := cluster %>% as.character()]

#' annotate cells based on annot table
#' 
#' @param obj seurat obj
#' @param subcluster_col string; 
annot_obj = function(obj, subcluster_col, annotdt){
  # remove unwanted subclusters
  cell_label = obj@meta.data[, subcluster_col]
  if(is.factor(cell_label)) cell_label = cell_label %>% unfactor() %>% as.character()
  obj = obj[, cell_label %in% annotdt$from]
  cell_label = obj@meta.data[, subcluster_col]
  if(is.factor(cell_label)) cell_label = cell_label %>% unfactor() %>% as.character()
  
  obj$cluster_group = annotdt[
    data.table(from = cell_label),
    on = .(from), to]
  DefaultAssay(obj) = "RNA"
  Idents(obj) = "cluster_group"
  return(obj)
}
```


# relabel cells by cluster annotation
```{r}
EC_ast = annot_obj(
  obj = EC_ast,
  subcluster_col = "integrated_snn_res.0.3",
  annotdt = clust_annot2[region == "EC", 
                       .(from = cluster,
                         to = cluster_group)]
)
BA20_ast = annot_obj(
  obj = BA20_ast,
  subcluster_col = "integrated_snn_res.0.3",
  annotdt = clust_annot2[region == "BA20", 
                       .(from = cluster,
                         to = cluster_group)]
)
BA46_ast = annot_obj(
  obj = BA46_ast,
  subcluster_col = "integrated_snn_res.0.3",
  annotdt = clust_annot2[region == "BA46", 
                       .(from = cluster,
                         to = cluster_group)]
)
V2_ast = annot_obj(
  obj = V2_ast,
  subcluster_col = "integrated_snn_res.0.3",
  annotdt = clust_annot2[region == "V2", 
                       .(from = cluster,
                         to = cluster_group)]
)
V1_ast = annot_obj(
  obj = V1_ast,
  subcluster_col = "integrated_snn_res.0.3",
  annotdt = clust_annot2[region == "V1", 
                       .(from = cluster,
                         to = cluster_group)]
)
```


# DE functions
```{r}
#' merge seurat obejcts
#'
#' @return a merged seurat obejct
merge_seurat = function(...){
  merged = Reduce(merge, list(...))
  return(merged)
}
```

```{r}
get_genes = function(obj){
  DefaultAssay(obj) = "RNA"
  return(rownames(obj))
}
genes = list(
  BA20 = get_genes(BA20_ast),
  BA46 = get_genes(BA46_ast),
  EC = get_genes(EC_ast),
  V1 = get_genes(V1_ast),
  V2 = get_genes(V2_ast)
)

union_genes = Reduce(union, genes)
inter_genes = Reduce(intersect, genes)

length(union_genes)
length(inter_genes)
writeLines(inter_genes, file.path(data_dir, "ast_intersect_seuratobject_genes.txt"))

load(file.path(data_dir, "Rdata", "ast_all.brain.regions_cluster.group_removed.Rdata"))
nrow(merged)
```


# DE: astH1 vs non-astH1
```{r}
merged = merge_seurat(
  EC_ast,
  BA20_ast,
  BA46_ast,
  V2_ast,
  V1_ast
)
merged = merged[inter_genes, ]
save(merged, file = file.path(data_dir, "Rdata", "ast_all.brain.regions_cluster.group_removed.Rdata"))

```


#Fig5A


This code is used to do correlation analysis based on HVGs


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load required packages
require(dendextend)
require(SNFtool)
require(data.table)
require(tidyverse)

# load required functions
source(file.path("..", "subcluster_functions.R"))

# directories
data_dir = file.path("..", "..", "data")
results_dir = file.path("..", "..", "results")
p_dir = file.path(results_dir, "plots", "all_brain_regions")

# load HVGs
ast_hvgs_dts = fread(file.path(results_dir, "ast_HVG_all.region.csv"))
ast_hvgs_dts.2 = fread(file.path(results_dir, "ast_HVG_all.region_removed.badclusters.csv"))

# load DE analysis result
load(file.path(data_dir, "gene_deg_all.Rdata"))
```


# a superset of HVGs
```{r}
# original HVGs used for clustering
ast_hvgs = ast_hvgs_dts[, hvgs] %>%
  unique() %>%
  sort()
# number of hvgs: 2855
ast_hvgs %>% length()
# this set of HVGs include some of the microglia genes because we have some contamination 
# clusters


# HVGs after remove containmination/donor specific subclusters
ast_hvgs.2 = ast_hvgs_dts.2[, hvgs] %>%
  unique() %>%
  sort()
# number of hvgs: 2942
ast_hvgs.2 %>% length()

# overlap: 0.9530931 
length(intersect(ast_hvgs, ast_hvgs.2))/length(ast_hvgs.2)

# venn of HVGs: before & after remove bad clusters
ggVennDiagram::ggVennDiagram(
  list(
    origin_HVG = ast_hvgs,
    new_HVG = ast_hvgs.2
  ), 
  label_alpha=0,
  label = "count"
  ) +
  scale_fill_gradient(low="white",high = "red")

ggVennDiagram(
  split(ast_hvgs_dts.2, ast_hvgs_dts.2$region) %>%
    map(., ~(.x$hvgs)),
  label_alpha=0,
  label = "count"
) +
  scale_fill_gradient(low="white",high = "red")
```


# gene rank correlation

## custome functions
```{r}
#' subset DE analysis list by selected genes
#' 
#' @param deg_dts list; a list of data.table that contains DE analysis results
#' @param features vector; a vector of gene names
#' 
#' @return a list of data.table for given genes
subset_deg = function(deg_dts, features){
  pblapply(deg_dts, function(deg_dt){
    return(deg_dt[gene %in% features, ])
  })
}

#' correlation analysis based on DE results (avg.logFC)
#' 
#' @param deg_dts list; a list of data.table that 
#' @param method string; method used for correlation analysis (default = "pearson")
#' 
#' @return a data.table contains correlation analysis results; 
cor_gene_ranks = function(deg_dts, method = "pearson"){
  # intersect of genes from all DE results
  inter_genes = Reduce(intersect, map(deg_dts, ~.x[, gene]))
  deg_dts = lapply(deg_dts, function(deg_dt){
    deg_dt[gene %in% inter_genes, ]
  })
  all_deg_dt = do.call(rbind, deg_dts)
  
  cors = pblapply(deg_dts, function(region_dt){
    subcluster_dts = split(region_dt, region_dt$cluster)
    dts = lapply(subcluster_dts, function(subcluster_dt){
      dt = subcluster_dt[all_deg_dt, on = .(gene)]
      # p-value
      dt[, avg.log2FC_cor_p := cor_test_results(
        x = avg_log2FC, y = i.avg_log2FC, 
        value = "p.value", method = method), 
        by = .(i.brain_region, i.cluster)]
      # R
      dt[, avg.log2FC_cor_R := cor_test_results(
        x = avg_log2FC, y = i.avg_log2FC, 
        value = "estimate", method = method), 
        by = .(i.brain_region, i.cluster)]
      return(
        dt[, .(subcluster1 = paste(brain_region, cluster, sep = ":"), 
               subcluster2 = paste(i.brain_region, i.cluster, sep = ":"), 
               avg.log2FC_cor_p, avg.log2FC_cor_R)] %>% unique()
             )
    })
    dts = do.call(rbind, dts)
    return(dts)
  })
  cors = do.call(rbind, cors)
  return(cors)
}

#' calculate affinity by Ayush
affinityCustom = function (diff, sigma = 0.5) 
{
  N <- nrow(diff)
  diff <- (diff + t(diff))/2
  diag(diff) <- 0
  sortedColumns <- as.matrix(t(apply(diff, 2, sort)))
  finiteMean <- function(x) {
    return(mean(x[is.finite(x)]))
  }
  # this line has been modified to remove [, 1:K + 1]
  means <- apply(sortedColumns, 1, finiteMean) + 
    .Machine$double.eps
  avg <- function(x, y) {
    return((x + y)/2)
  }
  Sig <- outer(means, means, avg)/3 * 2 + diff/3 + .Machine$double.eps
  Sig[Sig <= .Machine$double.eps] <- .Machine$double.eps
  densities <- dnorm(diff, 0, sigma * Sig, log = FALSE)
  W <- (densities + t(densities))/2
  return(W)
}


#' spectral clustering
#'
#' @param mtx matrix
#' @param k numeric; numer of clusters
#' 
#' @return a vector of clusters
spectral_clustering = function(mtx, k, seed = 77){
  set.seed(seed)

  # calculate distance matrix
  distM = mtx %>% dist2(., .) %>% .^(1/2)
  # calculate similarity matrix
  simM = affinityCustom(distM)
  # perform spectral clustering
  clustA = spectralClustering(simM, K = k)
  
  return(clustA)
}
```


## based on original HVGs used for clustering
```{r}
# DE analysis results
gene_deg.1 = subset_deg(
  deg_dts = gene_deg_all, 
  features = ast_hvgs
)
cor1 = cor_gene_ranks(
  deg_dts = gene_deg.1
)

# heatmap
cor1_mtx = cor1 %>% 
  dcast(subcluster1 ~ subcluster2, value.var = "avg.log2FC_cor_R") %>%
  as.data.frame() %>%
  column_to_rownames("subcluster1") %>% 
  as.matrix()

Heatmap(cor1_mtx)

# hierarchical clustering
set.seed(1234)
dend1 = hclust(dist(cor1_mtx))
plot(dend1)

# spectral clustering
sclust1 = spectral_clustering(cor1_mtx, k = 10)
Heatmap(cor1_mtx,
        row_split = sclust1,
        column_split = sclust1,
        cluster_rows = F,
        cluster_columns = F)
```

## based on HVGs after remove containmination/donor specific subclusters
```{r}
# DE analysis results
gene_deg.2 = subset_deg(
  deg_dts = gene_deg_all, 
  features = ast_hvgs.2
)
cor2 = cor_gene_ranks(
  deg_dts = gene_deg.2
)

# heatmap
cor2_mtx = cor2 %>% 
  dcast(subcluster1 ~ subcluster2, value.var = "avg.log2FC_cor_R") %>%
  as.data.frame() %>%
  column_to_rownames("subcluster1") %>% 
  as.matrix()


# hierarchical clustering
# require(dendextend)
# library(dendsort)
# set.seed(1234)
# dend2 = hclust(dist(cor2_mtx))
# row_dend = dendsort(dend2)
# col_dend = dendsort(hclust(dist(t(cor2_mtx))))
# plot(dend2)
# plot(row_dend)


# spectral clustering
sclust2 = spectral_clustering(cor2_mtx, k = 10)
Heatmap(cor2_mtx,
        row_split = sclust2,
        column_split = sclust2)

# cluster annotation
# Hs
clust_annot2 = data.table(
  region_cluster = c("V2:2", "V1:3", "V1:0", "BA20:0", "V2:0", "BA46:0"),
  cluster_group = "astH1"
)
clust_annot2 = rbind(clust_annot2, 
  data.table(
    region_cluster = c("EC:9", "V1:2", "EC:6", "EC:5", "EC:0"),
    cluster_group = "astH2"
))

# IMs
clust_annot2 = rbind(clust_annot2, 
  data.table(
    region_cluster = c("V1:8", "EC:8", "BA46:6"),
    cluster_group = "astIM1"
))
clust_annot2 = rbind(clust_annot2, 
  data.table(
    region_cluster = c("V2:1", "BA20:1", "BA46:1"),
    cluster_group = "astIM2"
))

clust_annot2 = rbind(clust_annot2, 
  data.table(
    region_cluster = c("BA46:4"),
    cluster_group = "astIM3"
))
clust_annot2 = rbind(clust_annot2, 
  data.table(
    region_cluster = c("V2:4", "BA20:2", "V1:6", "BA46:3"),
    cluster_group = "astIM4"
))
clust_annot2 = rbind(clust_annot2, 
  data.table(
    region_cluster = c("V1:1"),
    cluster_group = "astIM5"
))
clust_annot2 = rbind(clust_annot2, 
  data.table(
    region_cluster = c("EC:1", "BA20:5"),
    cluster_group = "astIM6"
))
# Rs
clust_annot2 = rbind(clust_annot2, 
  data.table(
    region_cluster = c("V2:5", "BA20:4", "V1:5", "EC:3"),
    cluster_group = "astR2"
))
clust_annot2 = rbind(clust_annot2, 
  data.table(
    region_cluster = c("EC:2", "V2:3", "BA20:3", "BA46:2", "V1:4"),
    cluster_group = "astR1"
))

# group
clust_annot2[, cluster_group := factor(cluster_group, levels = clust_annot2$cluster_group %>% unique())]
clust_annot2[, type := ifelse(grepl("astR", cluster_group), "astR",
                       ifelse(grepl("astH", cluster_group), "astH",
                              "IM"))]
clust_annot2[, color := ifelse(grepl("astR", cluster_group), "#C25757",
                       ifelse(grepl("astH", cluster_group), "#3A68AE",
                              "#E9C61D"))]
clust_annot2[, region := strsplit(region_cluster, ":") %>% 
               map(., ~(.x[1])) %>% 
               unlist()]
clust_annot2[, cluster := strsplit(region_cluster, ":") %>% 
               map(., ~(.x[2])) %>% 
               unlist()]

fwrite(clust_annot2, file = file.path(results_dir, "ast_cluster.annotation_HVG.after_Dec08.csv"))

clust_annot2 = fread(file.path(results_dir, "ast_cluster.annotation_HVG.after_Dec08.csv"))
nrow(clust_annot2)
dim(cor2_mtx)
cor2_mtx_order = clust_annot2[data.table(region_cluster = rownames(cor2_mtx)), 
                           on = .(region_cluster)]
require(circlize)
col_fun = colorRamp2(c(-1, 0, 1), c("#42B842", "white", "#862D84"))

colnames(cor2_mtx) = gsub("BA20", "ITG", colnames(cor2_mtx))
colnames(cor2_mtx) = gsub("BA46", "PFC", colnames(cor2_mtx))
colnames(cor2_mtx) = gsub("V2", "SVC", colnames(cor2_mtx))
colnames(cor2_mtx) = gsub("V1", "PVC", colnames(cor2_mtx))
rownames(cor2_mtx) = gsub("BA20", "ITG", rownames(cor2_mtx))
rownames(cor2_mtx) = gsub("BA46", "PFC", rownames(cor2_mtx))
rownames(cor2_mtx) = gsub("V2", "SVC", rownames(cor2_mtx))
rownames(cor2_mtx) = gsub("V1", "PVC", rownames(cor2_mtx))
write.csv(cor2_mtx, file = file.path(results_dir, "tables", "figure2_ast_cluster_correlation.csv"))

cor2_mtx_order$region_cluster = gsub("BA20", "ITG", cor2_mtx_order$region_cluster)
cor2_mtx_order$region_cluster = gsub("BA46", "PFC", cor2_mtx_order$region_cluster)
cor2_mtx_order$region_cluster = gsub("V2", "SVC", cor2_mtx_order$region_cluster)
cor2_mtx_order$region_cluster = gsub("V1", "PVC", cor2_mtx_order$region_cluster)

clust_annot2[, region_cluster_raw := region_cluster]
clust_annot2$region_cluster = gsub("BA20", "ITG", clust_annot2$region_cluster)
clust_annot2$region_cluster = gsub("BA46", "PFC", clust_annot2$region_cluster)
clust_annot2$region_cluster = gsub("V2", "SVC", clust_annot2$region_cluster)
clust_annot2$region_cluster = gsub("V1", "PVC", clust_annot2$region_cluster)
fwrite(clust_annot2, file.path(results_dir, "tables", "figure2_ast_cluster_group_annotation.csv"))


```



```{r}

# load cluster group annotation
clust_annot = fread(file.path("Data/Fig5_data/figure2_ast_cluster_group_annotation.csv"))
cluster_orders = c("astH1","astH2","astIM1","astIM2","astIM3","astIM4","astIM5","astIM6","astR2","astR1")
clust_annot[, cluster_group := factor(cluster_group, levels = cluster_orders)]

# load correlation analysis results
cor_mtx = read.csv(file.path("Data/Fig5_data/figure2_ast_cluster_correlation.csv"), row.names = 1) %>%
  as.matrix()
colnames(cor_mtx) = gsub("\\.", ":", colnames(cor_mtx))
cor_annot = clust_annot[data.table(region_cluster = rownames(cor_mtx)),
                         on = .(region_cluster)]

colors = c(
  "astH1" = "#2F79D2",
  "astH2" = "#7DCDF0", 
  "astIM1" = "#66c2a5",
  "astIM2" = "#abdda4", 
  "astIM3" = "#e6f598",
  "astIM4" = "#fee08b",
  "astIM5" = "#fdae61",
  "astIM6" = "#f46d43",
  "astR2" = "#D2514F", 
  "astR1" = "#C22023"
)

cor_annot[, color := colors[cluster_group]]
clust_annot[, color := colors[cluster_group]]

cluster_group_labels = as.list(cor_annot$cluster_group)
  
row_annot = rowAnnotation(
  df = cor_annot[, .(clusters = cluster_group)],
  col = list(clusters = colors),
  annotation_width = unit(c(1, 4), 'cm'),
  gap = unit(1, 'mm')
)
col_annot = HeatmapAnnotation(
  df = cor_annot[, .(clusters = cluster_group)],
  col = list(clusters = colors),
  annotation_width = unit(c(1, 4), 'cm'),
  gap = unit(1, 'mm'),
  show_legend = F
)

col_fun = colorRamp2(c(-1, 0, 1), c("blue", "white", "red"))
p_h = Heatmap(
  cor_mtx,
  # color scale
  col = col_fun, 
  # annotations
  right_annotation = row_annot,
  top_annotation = col_annot,
  # order of clusters (column & row of heatmap)
  row_order = clust_annot$region_cluster,
  column_order = clust_annot$region_cluster,
  # split heatmap by cluster group
  row_split = cor_annot$cluster_group,
  column_split = cor_annot$cluster_group,
  name = "correlation",
  row_names_side = "right",
  column_names_side = "top",
  row_title_side = "right",
  column_title_side = "top",
  row_title_gp = gpar(fontface = "bold", fontsize = 15),      
  column_title_gp = gpar(fontface = "bold",fontsize = 15),
  column_title_rot = 90,
  row_title_rot = 0,
  # black box around the cluster
  layer_fun = function(j, i, x, y, width, height, fill, slice_r, slice_c){
    v = pindex(cor_mtx, i, j)
    if(sum(v > 0.4)/length(v) > 0.4 & j[1] == i[1]) {
      grid.rect(gp = gpar(lwd = 3, fill = "transparent"))
    }
  }
)

png(file.path("~/Dropbox (Partners HealthCare)/Abbvie collaboration/Astrocytes scripts/Results/Figure_5/fig5a.png"),
    width = 700, height = 700)
draw(p_h)
dev.off()
```
# Figure 5b

proportion of CERAD cross brain region

```{r}
# proportion_bar = function(path_group, groups, regions, colors){
#   tlb = table(
#     path_group = paste0("Path ", path_group), 
#     group = groups,
#     region = regions
#   )
#   tlb = tlb %>%
#     as.data.frame() %>%
#     setDT()
#   setnames(tlb, "Freq", "ncells")
#   levels(tlb$region)[match("BA20",levels(tlb$region))] <- "ITG"
#   levels(tlb$region)[match("BA46",levels(tlb$region))] <- "PFC"
#   tlb[, region := factor(region, levels = c("EC", "ITG", "PFC", "V2", "V1"))]
#   tlb[, total_cells := sum(ncells), by = .(region, path_group)]
#   tlb[, pct := ncells/total_cells]
#   
#   tlb[, g_ncells := sum(ncells), by = .(group, path_group)]
#   tlb[, g_sum_cells := sum(ncells)]
#   tlb[, g_pct := g_ncells/g_sum_cells]
#   
#   tlb[, binomial_p := ifelse(g_pct == 0, 100, 
#           binom.test(ncells, total_cells, p = g_pct,
#             alternative = "two.sided",
#             conf.level = 0.95)$p.value), by = rownames(tlb)]
#   
#   tlb[, binomial_p_label := ifelse(binomial_p == 100, "",
#                               ifelse(binomial_p > 0.05, "NS",
#                               ifelse(binomial_p > 0.01, "*",
#                                ifelse(binomial_p > 0.001, "**", 
#                                 ifelse(binomial_p <= 0.001, "***",
#                                        "")))))]
# 
#   tlb[, group := factor(group, levels = names(colors))]
#   
#   p = ggplot(tlb, 
#        aes(x = path_group, y = pct, fill = group)) +
#     geom_bar(stat = "identity", position="fill", color = "black")  +
#     # geom_text(data = subset(tlb, ncells!= 0),
#     #           aes(label = paste0(round(pct * 100, 1),"%")), 
#     #           position = position_stack(vjust = 0.5), size = 3,
#     #           angle = 90) +
#     scale_y_continuous(labels = scales::percent) +
#     facet_grid(. ~ region, space = "free_x", scales = "free_x") +
#     my_border_theme() +
#     theme(legend.position="top") +
#     theme(axis.text.x=element_text(size= 10, angle = 90)) +
#     labs(y = "Proportion", x = "Brain regions") +
#     scale_fill_manual(
#       name = "ast clusters",
#       values = colors
#     ) 
#   return(p)
# }
# 
# p_IMHR = proportion_bar(
#   path_group = cluster_groups_subset$Path..Group., 
#   groups = gsub("IM\\d$", "IM", cluster_groups_subset$cluster_group),
#   regions = cluster_groups_subset$Unified_region, 
#   colors = c(
#     "astH1" = "#2F79D2", # "#3A8CAF", 
#     "astH2" = "#7DCDF0", #"#66c2a5",
#     "astIM" = "#FFFDA2",
#     "astR2" = "#D2514F", #"#EB6532",
#     "astR1" = "#C22023"
#   )
# )
# p_IMHR
# ggsave(p_IMHR, filename = file.path(p_dir, "fig4b_p_IMHR_05022022.png"), 
#        height = 5, width = 8)
```


#Fig 5b
```{r}
load("~/Dropbox (Partners HealthCare)/Abbvie collaboration/Astrocytes scripts/Data/Fig5_data/merged_4c.rdata")

merged_4c <- merged_new

astHR = merged_4c[, merged_4c$cluster_group %in% c("astH1", "astH2", "astR1", "astR2")]

avge = AverageExpression(merged_4c, assays = "RNA", features = rownames(merged_4c)) %>%
  .[["RNA"]]
avge = avge %>% 
  as.data.frame() %>%
  rownames_to_column("gene_name") %>%
  as.data.table() %>%
  melt(id.vars = c("gene_name"), variable.name = "clusters", 
       value.name = "clusters_ave_exp") %>%
  .[, ave_exp_z := scale(clusters_ave_exp), by = .(gene_name)]
avge = avge %>% 
  dcast(gene_name ~ clusters, value.var = "ave_exp_z")


zcut = 0

pct = 0.5

astH1 = fread(file.path("Data/fig5_data/astH1_de_12112021.csv"))
astH2 = fread(file.path("Data/fig5_data/astH2_de_12112021.csv"))
astH1vsH2 = fread(file.path("Data/fig5_data/astH1vsastH2_de_12112021.csv"))
astR1 = fread(file.path("Data/fig5_data/astR1_de_12112021.csv"))
astR2 = fread(file.path("Data/fig5_data/astR2_de_12112021.csv"))
astR1vsR2 = fread(file.path("Data/fig5_data/astR1vsastR2_de_12112021.csv"))

# H1,H2 common markers
astH1H2_c = Reduce(intersect,
                   list(
                     astH1[pct.1 > pct, gene_name],
                     astH2[pct.1 > pct, gene_name],
                     avge[astH1 > zcut & astH2 > zcut & astR1 < -zcut & astR2 < -zcut, gene_name]
                   )
)

# R1,R2 common markers
astR1R2_c = Reduce(intersect,
                   list(
                     astR1[pct.1 > pct, gene_name],
                     astR2[pct.1 > pct, gene_name],
                     avge[astR1 > zcut & astR2 > zcut & astH1 < -zcut & astH2 < -zcut, gene_name]
                   )
)

# H1 markers
astH1_gs = Reduce(intersect,
                  list(
                    astH1[pct.1 > pct, gene_name],
                    avge[astH1 > zcut & astH2 < 0 & astR1 < 0 & astR2 < 0, gene_name]
                  )
)

# H2 genes
astH2_gs = Reduce(intersect,
                  list(
                    astH2[pct.1 > pct, gene_name],
                    avge[astH2 > zcut & astH1 < 0 & astR1 < 0 & astR2 < 0, gene_name]
                  )
)

# R1 markers
astR1_gs = Reduce(intersect,
                  list(
                    astR1[pct.1 > pct, gene_name],
                    avge[astR1 > zcut & astH2 < 0 & astH1 < 0 & astR2 < 0, gene_name]
                  )
)

# R2 genes
astR2_gs = Reduce(intersect,
                  list(
                    astR2[pct.1 > pct, gene_name],
                    avge[astR2 > zcut & astH1 < 0 & astR1 < 0 & astH2 < 0, gene_name]
                  )
)


fig3c_dt = rbind(
  data.table(
    gene = astH1_gs,
    group = "astH1"
  ),
  data.table(
    gene = astH2_gs,
    group = "astH2"
  ),
  data.table(
    gene = astR1_gs,
    group = "astR1"
  ),
  data.table(
    gene = astR2_gs,
    group = "astR2"
  ),
  data.table(
    gene = astH1H2_c,
    group = "astH1/H2 common"
  ),
  data.table(
    gene = astR1R2_c,
    group = "astR1/R2 common"
  )
)
fig3c_dt[, group := factor(group, levels = c("astH1", "astH2", "astH1/H2 common",
                                             "astR1", "astR2", "astR1/R2 common"))]


fig5bgenes = c("GJA1","CTNNA2", "EFNA5", "FGFR2", "HPSE2", "IGF1R", "KCNMA1", "MBNL2", "NFE2L2", "NRF1", "NRG2", "AKAP9", "BCL2", "ETNPPL","GLUD1","MT-ATP6", "MT-CO1", "MT-CO2", "MT-CO3", "MT-ND1", "MT-ND3", "RBMS3", "ALDH1L1", "CD47","EGFR", "EGLN3", "ERBB4", "GLUL", "GRM3", "HIF3A", "LRP1","MEGF10", "NFIA","NFIB", "NRXN1","NTRK3","PON2","PTN","RYR3","SLC1A2","SLC1A3","SYNE1","ADAMTSL3","AQP1","CAMK2D","GAB1","HIFIA","IRS2","LONP2","MAP4K3","SOD2","SORT1","EZR","CLU","GAREM1","GRIA1","LRIG1","MAPK1","MARK3","PRKCA","TRA2A","TRAK1","ACTB","APP","AQP4","CD44","CRYAB","ECE1","FGFR1","GFAP","HSPB1","HSPB8","LAMA1","MAOB","MAP1B","MAPT","MT1E","MT2A","PRDX6","S100B","VCAN")
fig5b_dt <- fig3c_dt[gene %in% fig5bgenes,]
fig5b_dt<-add_row(fig5b_dt, gene = "GJA1", group = "astH1/H2 common")

obj = merged_4c
features_dt = fig3c_dt
sel_group = "astH1"

do_dotplot = function(obj, features_dt, sel_group = NULL, h = 20, w = 7){
  DefaultAssay(obj) = "RNA"
  
  if(!is.null(sel_group)){
    features_dt = features_dt[group %in% sel_group,]
  }
  avge = DotPlot(obj, features = features_dt$gene)$data %>%
    as.data.frame() %>%
    setDT()
  avge = avge[id %in% c("astH1", "astH2", "astR1", "astR2"), ]
  avge[, group := features_dt[avge, on = c("gene" = "features.plot"), group]]
  avge[, group := factor(group, levels = c(
    "astH1", "astH2", "astH1/H2 common",
    "astR1", "astR2", "astR1/R2 common"
  ))]
  avge[, sugroup := features_dt[avge, on = c("gene" = "features.plot"), group]]
  avge[, features.plot := features.plot %>% as.character()]
  avge[, features.plot := factor(features.plot, 
      levels = sort(features.plot %>% unique(), decreasing = T))]
  
   pd = ggplot(avge, aes(
     x = id, y = features.plot, 
     color = avg.exp.scaled, size = pct.exp)
    ) +
    geom_point() +
    scale_color_gradient2(
        name = "Expression \n z-score", 
        low="#2F79D2",
        mid="white",
        high="#C22023", na.value = "gray", 
        midpoint = 0,
        limits = c(-2.1, 2.5)
      ) +
     scale_size(
        range = c(1, 6),
        breaks = c(25, 50, 75, 100)
     ) +
     facet_grid(group ~ ., space = "free", scales = "free") +
     my_border_theme()+ 
     labs(y = "Genes")
  return(list(p = pd, data = avge))
}

# changing the data.table to remove MT genes:

fig5b_dt = fread("~/Dropbox (Partners HealthCare)/Abbvie collaboration/Astrocytes scripts/Scripts-git/ad-progression-study/genes_fig5c.csv")

fig5b_H = do_dotplot(
  merged_4c, features_dt = fig5b_dt,
  sel_group = c("astH1", "astH2", "astH1/H2 common")
)
fig5b_R = do_dotplot(
  merged_4c, features_dt = fig5b_dt,
  sel_group = c("astR1", "astR2", "astR1/R2 common")
)

fig5b_H$data$avg.exp.scaled %>% range()
fig5b_R$data$avg.exp.scaled %>% range()

require(ggpubr)
fig5b = ggarrange(
  fig5b_H$p + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(angle = 15)),
  fig5b_R$p + theme(axis.title.y = element_blank() + theme(axis.text.x = element_text(15)),
                    axis.title.x = element_blank()) + theme(axis.text.x = element_text(angle = 15)),
  ncol = 2, nrow = 1,
  align = "hv",
  common.legend = T,
  legend = "right"
)
#fwrite(fig3c_dt, file.path(results_dir, "ast_group_zscore_markers.csv"))

# p_fig3c = do_dotplot(astHR, features_dt = fig3c_dt)

fig5b

ggsave(
  fig5b, 
  filename = file.path("~/Dropbox (Partners HealthCare)/Abbvie collaboration/Astrocytes scripts/Results/Figure_5/figure5c.png"), 
  height = 10, width = 8,
  limitsize = FALSE
)

#fwrite(fig5b_dt, "data.csv")
#fig3c_dt = fread(file.path(data_dir, "Alberto_ast_group_markers.csv"))

p_astH1 = do_dotplot(merged_4c, features_dt = fig5b_dt, sel_group = "astH1")
p_astH2 = do_dotplot(merged_4c, features_dt = fig5b_dt, sel_group = "astH2")
p_astR1 = do_dotplot(merged_4c, features_dt = fig5b_dt, sel_group = "astR1", h = 60)
p_astR2 = do_dotplot(merged_4c, features_dt = fig5b_dt, sel_group = "astR2")
p_astR1R1c = do_dotplot(merged_4c, features_dt = fig5b_dt, sel_group = "astR1/R2 common", h = 60)
p_astH1H2c = do_dotplot(merged_4c, features_dt = fig5b_dt, sel_group = "astH1/H2 common", h = 60)
```

#Fig 5d

```{r}
ncell_dt = fread(file.path(data_dir, "proportion_cells.csv"))
ncell_dt[, ast_cluster := ifelse(is.na(group), "astIM", group)]
ncell_dt[, ast_cluster := factor(ast_cluster, labels = c("astH", "astH", "astIM", "astR", "astR"))]
ncell_dt2 = ncell_dt[, .(ncells_ast_cluster = sum(ncells)), by = .(ast_cluster, DonorID, region)]
ncell_dt2[, region := factor(region, levels = c("EC", "ITG", "PFC", "V2", "V1"))]

ncell_dt2[, ncell_sample := sum(ncells_ast_cluster), by = .(DonorID, region)]
ncell_dt2[, pct_sample := ncells_ast_cluster/ncell_sample]

ncell_dt2[ncell_dt, on = .(DonorID), path_group := i.path_group]

ncell_dt2[, mean_pct_trend := mean(pct_sample, na.rm = T), by = .(path_group, region, ast_cluster)]
ncell_dt2[, mean_id := paste(region, path_group, ast_cluster, sep = "_")]
ncell_dt2[, sd_pct_trend := sd(pct_sample, na.rm = T), by = .(path_group, region, ast_cluster)]
ncell_path = copy(ncell_dt2)
ncell_path[, region := factor(region, levels = c("EC", "ITG", "PFC", "V2", "V1"))]
ncell_path[, path_mean_pct_trend := mean(pct_sample, na.rm = T), by = .(path_group, ast_cluster)]
ncell_path[, path_sd_pct_trend := sd(pct_sample, na.rm = T), by = .(path_group, ast_cluster)]
ncell_path[, path_median_pct_trend := mean(pct_sample, na.rm = T), by = .(path_group, ast_cluster)]
ncell_path = ncell_path[, .(path_group, region, ast_cluster, path_mean_pct_trend, path_sd_pct_trend, path_median_pct_trend)] %>% unique()
ncell_region = copy(ncell_dt2)
ncell_region[, region_mean_pct_trend := mean(pct_sample, na.rm = T), by = .(region, ast_cluster)]
ncell_region[, region_median_pct_trend := median(pct_sample, na.rm = T), by = .(region, ast_cluster)]
ncell_region[, region := factor(region, levels = c("EC", "ITG", "PFC", "V2", "V1"))]
ncell_region[, region_sd_pct_trend := sd(pct_sample, na.rm = T), by = .(region, ast_cluster)]
ncell_region = ncell_region[, .(path_group, region, ast_cluster, region_mean_pct_trend, region_sd_pct_trend, region_median_pct_trend)] %>% unique()

ncell_path<- ncell_path[,pathology_groups := gsub("Path ", "", path_group)]
ncell_dt2<- ncell_dt2[,pathology_groups := gsub("Path ", "", path_group)]

ggplot(
  unique(ncell_path[, .(pathology_groups, path_mean_pct_trend, ast_cluster)])
) +
  geom_bar(aes(x = pathology_groups, y = path_mean_pct_trend, group = ast_cluster,
                fill = ast_cluster), stat = "identity", width = 0.4) +
  facet_grid(ast_cluster~.) +
  my_border_theme() +
  labs(x = "Pathology stage", y = "Proportion") +
  theme(legend.position = "top") +
  theme(legend.title = element_blank())+
  geom_jitter(
    data = ncell_dt2,
    aes(x = pathology_groups, y = pct_sample),
    size = 0.5, width = 0.1, color = "black") +
  scale_fill_manual(values = c("lightblue", "orange", "red"))

ggsave(file.path(rdir, "AD progression", "proportion_path_bar.png"), width = 3)


ggplot(
  unique(ncell_region[, .(region, region_median_pct_trend, ast_cluster)])
) +
  
  geom_bar(aes(x = region, y = region_median_pct_trend, group = ast_cluster,
                fill = ast_cluster), stat = "identity", width = 0.4) +
  theme(legend.title = element_blank())+
  geom_jitter(
    data = ncell_dt2,
    aes(x = region, y = pct_sample),
    size = 0.5, width = 0.1, color = "black") +
  facet_grid(ast_cluster~.) +
  my_border_theme() +
  labs(x = "Regions", y = "Proportion") +
  theme(legend.position = "top") +
  scale_fill_manual(values = c("lightblue", "orange", "red"))
ggsave(file.path(rdir, "AD progression", "proportion_region_bar.png"), width = 3)
```

